### 2008-05-01 documents/WCSS2008Paper/paper.tex
1) Added some missing papers to papers folder and created documents folder. 2) Created skeleton of presentation (to be found in documents folder), using Powerpoint 2003 (my understanding is that both Mac as well as Win people are OK with this choice). 3) Created skeleton of the Latex paper: we need the paper to submit for WCSS2008 at some point and they require Springer formatting which is unpleasant to work with in Word. Included bibtex bibliography.

\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
\usepackage{lscape}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage{color}
\usepackage{tikz}
\usepackage{amssymb}
\usepackage{amsmath}

\urldef{\mailsg}\path|{rax222}@gmu.edu|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter

\title{Simulating Financial Markets using MASON Framework\thanks{Prepared for the 2nd World Congress on Social Simulation, GMU, Fairfax, 14--17 July, 2008.}}

\titlerunning{Financial Markets using MASON}
\authorrunning{R. Axtell et al.}

\author{Robert Axtell$^{a}$ \and CSS 739 Class Project Team}

\institute{
4400 University Drive, Fairfax VA 22030 \\
\mailss \\
\url{http://www.si.edu}}

\institute{$^{a}$Center for Social Complexity, George Mason University, USA \\
\mailsg}

\date{\today}
\maketitle
\begin{abstract}
AAA
\end{abstract}

\keywords{Agent-based Modeling, Computational Social Science, Financial Markets}

\section{Motivation and Objectives}

\section{Platform Architecture}

\begin{figure}[htb]
\centering
\includegraphics[width=1.0\textwidth]{../graphics/masterClassDiagram.jpg}
\caption{High-level UML class diagram of the main components and relations in the FinancialMarketModel, including the main attributes of Players and OrderBooks. Agent classes (light blue) inherit from the MASON \texttt{Steppable} interface while the master class is implementation of MASON's \texttt{SimState}.}
\label{fig:general_uml}
\end{figure}

\section{Verification of Correctness}

\section{Overview of Implemented Models}

\begin{itemize}
	\item \cite{farmer2003}
	\item \cite{lamba2007}
	\item \cite{westerhoff2004}
	\item \cite{cont2006}
\end{itemize}

\bibliographystyle{plainnat}
\bibliography{master}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The end.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2008-05-01 src/gui/FinancialModelWithUI.java
1) Added volumes and average trade prices to ordersBooks, reporters and graphs 2) Modified Reporter to produce files in formats used by Data Analysis group

@@ -35,6 +35,10 @@ public class FinancialModelWithUI extends GUIState {
 	public TimeSeriesChartGenerator acfChart;
 
 	public JFrame acfFrame;
+	
+	public TimeSeriesChartGenerator volumeChart;
+
+	public JFrame volumeFrame;
 
 	public HistogramGenerator returnHist;
 
@@ -184,6 +188,19 @@ public class FinancialModelWithUI extends GUIState {
 		returnFrame.getContentPane().add(returnChart, BorderLayout.CENTER);
 		returnFrame.pack();
 		c.registerFrame(returnFrame);
+		
+		volumeChart = new TimeSeriesChartGenerator();
+		volumeChart.setTitle("Volumes plot");
+		volumeChart.setDomainAxisLabel("Step");
+		volumeChart.setRangeAxisLabel("Volume");
+		for (int a = 0; a < myModel.parameterMap.get("numAssets"); a++) {
+			volumeChart.addSeries(myReporter.volumeSeries.get(a), null);
+		}
+		volumeFrame = volumeChart.createFrame(this);
+		volumeFrame.getContentPane().setLayout(new BorderLayout());
+		volumeFrame.getContentPane().add(volumeChart, BorderLayout.CENTER);
+		volumeFrame.pack();
+		c.registerFrame(volumeFrame);
 
 		acfChart = new TimeSeriesChartGenerator();
 		acfChart.setTitle("Autocorrelation of returns");

### 2008-05-01 src/gui/GUIReporter.java

@@ -26,13 +26,17 @@ public class GUIReporter implements Steppable {
 	public ArrayList<XYSeries> acfAbsReturnsSeries = new ArrayList<XYSeries>();
 
 	public ArrayList<XYSeries> acfReturnsSeries = new ArrayList<XYSeries>();
+	
+	public ArrayList<XYSeries> volumeSeries = new ArrayList<XYSeries>();
 
 	public ArrayList<Vector<Double>> bidPriceMemory = new ArrayList<Vector<Double>>();
 	
 	public ArrayList<Vector<Double>> askPriceMemory = new ArrayList<Vector<Double>>();
 
 	public ArrayList<Vector<Double>> returnMemory = new ArrayList<Vector<Double>>();
-
+	
+	public ArrayList<Vector<Double>> volumeMemory = new ArrayList<Vector<Double>>();
+	
 	public int NumViewable = 10000; // also, the #/data points used in
 
 	// calculating acf
@@ -54,7 +58,7 @@ public class GUIReporter implements Steppable {
 			askPriceSeries.add(new XYSeries("Ask price for asset " + a));
 			returnSeries.add(new XYSeries("Returns for asset " + a));
 			absReturnSeries.add(new XYSeries("Absolute returns for asset " + a));
-
+			volumeSeries.add(new XYSeries("Trade volumes for asset " + a));
 			acfAbsReturnsSeries.add(new XYSeries("ACF of absolute returns for asset " + a));
 			acfReturnsSeries.add(new XYSeries("ACF of returns for asset " + a));
 
@@ -62,6 +66,7 @@ public class GUIReporter implements Steppable {
 			bidPriceSeries.get(a).setMaximumItemCount(NumViewable);
 			askPriceSeries.get(a).setMaximumItemCount(NumViewable);
 			returnSeries.get(a).setMaximumItemCount(NumViewable);
+			volumeSeries.get(a).setMaximumItemCount(NumViewable);
 			absReturnSeries.get(a).setMaximumItemCount(NumViewable);
 			// initialize
 			for (int i = 1; i < ACFViewable; i++) {
@@ -72,6 +77,7 @@ public class GUIReporter implements Steppable {
 			bidPriceMemory.add(new Vector<Double>());
 			askPriceMemory.add(new Vector<Double>());
 			returnMemory.add(new Vector<Double>());
+			volumeMemory.add(new Vector<Double>());
 		}
 
[…]

### 2008-05-01 src/model/market/Market.java

@@ -84,4 +84,14 @@ public class Market implements Steppable {
 		return orderBooks.get(i).getRandomComponent();
 	}
 
+	public double getAverageTradePriceForAsset(int i) {
+		// TODO Auto-generated method stub
+		return orderBooks.get(i).getAverageTradePrice();
+	}
+
+	public double getVolumeForAsset(int i) {
+		// TODO Auto-generated method stub
+		return orderBooks.get(i).getVolume();
+	}
+
 }

### 2008-05-01 src/model/market/books/ContBook.java

@@ -22,10 +22,22 @@ public class ContBook implements OrderBook {
 	
 	// external signal to update threshold
 	public double epsilon_t;
-
 	
+	public double currentDemand = 0;
+	
+	public double currentSupply = 0;
+	
+	public double oldVolume = 0;
+
+	public double averageTradePrice = 0;
 
 	public void cleanup() {
+		
+
+		oldVolume = Math.min(currentDemand, currentSupply);
+		
+		averageTradePrice = oldVolume*price_t;
+		
 		// calculate return rate; the excess demand is already known
 		// as it is called by agents during order generation phase
 		returnRate_t = this.priceImpact(excessDemand / myWorld.agentList.size());
@@ -35,6 +47,10 @@ public class ContBook implements OrderBook {
 		excessDemand = 0.0;
 		// set shock for next period
 		epsilon_t = myWorld.parameterMap.get("D") * myWorld.random.nextGaussian();
+		
+		currentDemand = 0;
+		
+		currentSupply = 0;
 	}
 
 
@@ -47,8 +63,10 @@ public class ContBook implements OrderBook {
 		
 		if (order.type.equals(OrderType.PURCHASE)) {
 			excessDemand += order.quantity;
+			currentDemand++;
 		} else {
 			excessDemand -= order.quantity;
+			currentSupply++;
 		}
 		
 		return true;
@@ -118,6 +136,20 @@ public class ContBook implements OrderBook {
 	public double getRandomComponent() {
 		// TODO Auto-generated method stub
 		return epsilon_t;
+	}
+
+
+
+	public double getVolume() {
+		// TODO Auto-generated method stub
+		return this.oldVolume;
+	}
+
+
+
+	public double getAverageTradePrice() {
+		// TODO Auto-generated method stub
+		return  this.averageTradePrice;
 	}
 	
 	

### 2008-05-01 src/model/market/books/DoubleAuctionOrderBook.java

@@ -21,16 +21,27 @@ public class DoubleAuctionOrderBook implements OrderBook {
 	protected FinancialModel myWorld;
 
 	protected SortedSet<LimitOrder> buyOrders;
+
 	protected SortedSet<LimitOrder> sellOrders;
 
 	protected int myID;
 
 	public double returnRate_t = 0;
+
 	public double price_t = 1;
-	
-	// set to true if price as stored in the order book are the log of actual price;
+
+	// set to true if price as stored in the order book are the log of actual
+	// price;
 	// this implies negative prices are perfectly acceptable
 	public boolean logPricing = false;
+
+	public double volumeTraded = 0;
+
+	public double valueTraded = 0;
+	
+	public double oldVolumeTraded = 0;
+
+	public double oldValueTraded = 0;
 
 	public DoubleAuctionOrderBook() {
 		super();
@@ -45,8 +56,8 @@ public class DoubleAuctionOrderBook implements OrderBook {
 		}
 
 		// give it a new, unique transaction ID within the orderbook
-//		order.transactionID.set(nextTransactionID++);
-		
+		// order.transactionID.set(nextTransactionID++);
+
 		if (order.type == OrderType.PURCHASE) {
 			buyOrders.add(order);
 		} else {
@@ -78,41 +89,45 @@ public class DoubleAuctionOrderBook implements OrderBook {
 		double currentTime = myWorld.schedule.getTime();
 		int origQuant = quantity;
 		double totalPrice = 0.0;
-        
+
 		// The operation is the same regardless of whether we're buying or
 		// selling. We just need to choose the right orders to work on.
 		SortedSet<LimitOrder> orders;
 		if (type == OrderType.PURCHASE) {
-			orders=sellOrders;
+			orders = sellOrders;
 		} else {
-			orders=buyOrders;
+			orders = buyOrders;
 		}
 
 		while ((quantity > 0) && (!orders.isEmpty())) {
 
 			LimitOrder lo = orders.first();
 
- 			// ensure this order hasn't expired
+			// ensure this order hasn't expired
 			if (lo.expirationTime < currentTime) {
 				continue;
 			}
- 			
-			int curQuantity=0;
- 			if (lo.quantityPending() >= quantity) {
- 				curQuantity = quantity;
- 			} else {
- 				curQuantity = quantity - lo.quantityPending();
- 			}
[…]

### 2008-05-01 src/model/market/books/LSBook.java

@@ -118,4 +118,14 @@ public class LSBook implements OrderBook {
 		return epsilon_t;
 	}
 
+	public double getVolume() {
+		// TODO Auto-generated method stub
+		return 0;
+	}
+
+	public double getAverageTradePrice() {
+		// TODO Auto-generated method stub
+		return 0;
+	}
+
 }

### 2008-05-01 src/model/market/books/OrderBook.java

@@ -68,4 +68,9 @@ public interface OrderBook {
 	public void setMyID(int a);
 
 	public double getRandomComponent();
+	
+	public double getVolume();
+	
+	public double getAverageTradePrice();
+	
 }

### 2008-05-01 src/support/Reporter.java

@@ -29,11 +29,13 @@ public class Reporter implements Steppable {
 			
 			
 			String temp = "";
-			temp = temp + "runID" + "	";
-			temp = temp + "time" + "	";
-			temp = temp + "asset" + "	";
-			temp = temp + "askPrice" + "	";
-			temp = temp + "bidPrice" + "	";
+			temp = temp + "runID" + ";";
+			temp = temp + "time" + ";";
+			temp = temp + "asset" + ";";
+			temp = temp + "closeAskPrice" + ";";
+			temp = temp + "closeBidPrice" + ";";
+			temp = temp + "averageTradePrice" + ";";
+			temp = temp + "volume" + ";";
 			temp = temp + "return" ;
 			
 
@@ -64,12 +66,14 @@ public class Reporter implements Steppable {
 			for (int asset =0; asset < myModel.parameterMap.get("numAssets"); asset++) {
 			
 			String temp = "";
-			temp = temp + myModel.runID + "	";
-			temp = temp + myModel.schedule.getTime() + "	";
-			temp = temp + myModel.myMarket.getAskPriceForAsset(asset)+ "	";
-			temp = temp + myModel.myMarket.getBidPriceForAsset(asset)+ "	";
-			temp = temp + myModel.myMarket.getReturnRateForAsset(asset);
-		
+			temp = temp + myModel.runID + ";";
+			temp = temp + myModel.schedule.getTime() + ";";
+			temp = temp + myModel.myMarket.getAskPriceForAsset(asset)+ ";";
+			temp = temp + myModel.myMarket.getBidPriceForAsset(asset)+ ";";
+			temp = temp + myModel.myMarket.getAverageTradePriceForAsset(asset) + ";";
+			temp = temp + myModel.myMarket.getVolumeForAsset(asset) + ";";
+			temp = temp + myModel.myMarket.getReturnRateForAsset(asset) + ";";
+			
 			outPrices.write(temp);
 			outPrices.newLine();			
 

### 2008-05-01 documents/WCSS2008Paper/paper.tex
1) Included a mash-up of graphic outputs for article. 2) We have broken the acf functions somewhere, a bug to be corrected

@@
 
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth,height=0.8\textwidth]{../graphics/pricesPlot.pdf}}} \quad
+      \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/returnPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Orderbook shape.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/orderbooks.pdf}}} \quad
+      \subfigure[Trade volumes.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/volumesPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Returns histogram.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/returnsHistogram.pdf}}} \quad
+      \subfigure[Autocorrelation of returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/acfReturns.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation for default Farmer's parametrization.}
+    \label{fig:sampleDynamics}
+  \end{center}
+\end{figure}
+
 \section{Verification of Correctness}

### 2008-05-01 src/gui/GUIReporter.java

@@ -37,7 +37,7 @@ public class GUIReporter implements Steppable {
 	
 	public ArrayList<Vector<Double>> volumeMemory = new ArrayList<Vector<Double>>();
 	
-	public int NumViewable = 10000; // also, the #/data points used in
+	public int NumViewable = 6000; // also, the #/data points used in
 
 	// calculating acf
 

### 2008-05-03 src/model/market/Market.java
DoubleAuctionOrderBook supports sending shock signals. Cleanup and initialization of epsilon_t.

@@ -28,10 +28,11 @@ public class Market implements Steppable {
 	// constructor
 	public Market(FinancialModel myWorld) {
 		this.myWorld = myWorld;
-
+/*
 		for (int o = 0; o < myWorld.parameterMap.get("numAssets"); o++) {
 
 		}
+*/
 
 	}
 
@@ -73,7 +74,6 @@ public class Market implements Steppable {
 		try {
 			this.orderBooks.get(asset).executeMarketOrder(newType, amount);
 		} catch (LiquidityException e) {
-			// TODO Auto-generated catch block
 			// TODO: this should be left for the caller to catch
 			 e.printStackTrace();
 		}
@@ -85,12 +85,10 @@ public class Market implements Steppable {
 	}
 
 	public double getAverageTradePriceForAsset(int i) {
-		// TODO Auto-generated method stub
 		return orderBooks.get(i).getAverageTradePrice();
 	}
 
 	public double getVolumeForAsset(int i) {
-		// TODO Auto-generated method stub
 		return orderBooks.get(i).getVolume();
 	}
 

### 2008-05-03 src/model/market/books/ContBook.java

@@ -21,7 +21,7 @@ public class ContBook implements OrderBook {
 	public double returnRate_t = 0;
 	
 	// external signal to update threshold
-	public double epsilon_t;
+	public double epsilon_t = 0.1;
 	
 	public double currentDemand = 0;
 	
@@ -79,8 +79,6 @@ public class ContBook implements OrderBook {
 		// return Math.atan2(d, myWorld.lambda);
 		// return d / myWorld.lambda;
 	}
-
-	
 
 	public void setMyWorld(FinancialModel myWorld) {
 		this.myWorld = myWorld;

### 2008-05-03 src/model/market/books/DoubleAuctionOrderBook.java

@@ -36,12 +36,11 @@ public class DoubleAuctionOrderBook implements OrderBook {
 	public boolean logPricing = false;
 
 	public double volumeTraded = 0;
-
 	public double valueTraded = 0;
-	
 	public double oldVolumeTraded = 0;
-
 	public double oldValueTraded = 0;
+	
+    public double epsilon_t = 0.01;
 
 	public DoubleAuctionOrderBook() {
 		super();
@@ -121,7 +120,7 @@ public class DoubleAuctionOrderBook implements OrderBook {
 			lo.quantityExecuted.addAndGet(curQuantity);
 			// Update the market order
 			totalPrice += lo.pricePerUnit * curQuantity;
-
+            // Update aggregate value/volume traded
 			this.valueTraded += lo.pricePerUnit * curQuantity;
 			this.volumeTraded += curQuantity;
 
@@ -143,7 +142,8 @@ public class DoubleAuctionOrderBook implements OrderBook {
 
 		if (buyOrders.isEmpty()) {
 			// TODO: this should probably change to price_t
-			return myWorld.parameterMap.get("initialPrice");
+			//return myWorld.parameterMap.get("initialPrice");
+			return price_t;
 		} else {
 			return buyOrders.first().pricePerUnit;
 		}
@@ -154,7 +154,8 @@ public class DoubleAuctionOrderBook implements OrderBook {
 
 		if (sellOrders.isEmpty()) {
 			// TODO: this should probably change to price_t
-			return myWorld.parameterMap.get("initialPrice");
+			//return myWorld.parameterMap.get("initialPrice");
+			return price_t;
 		} else {
 			return sellOrders.first().pricePerUnit;
 		}
@@ -227,7 +228,9 @@ public class DoubleAuctionOrderBook implements OrderBook {
 		this.valueTraded = 0;
 		this.oldVolumeTraded = this.volumeTraded;
 		this.volumeTraded = 0;
-		
+
+		// set shock for next period
+		epsilon_t = myWorld.parameterMap.get("D") * myWorld.random.nextGaussian();
 	}
 
 	// returns an array with an entry of the price for each unit of limit order
@@ -295,17 +298,14 @@ public class DoubleAuctionOrderBook implements OrderBook {
 	}
 
 	public double getRandomComponent() {
-		// TODO Auto-generated method stub
-		return 0;
+		return epsilon_t;
 	}
 
 	public double getVolume() {
[…]

### 2008-05-03 src/model/market/books/LSBook.java

@@ -21,7 +21,7 @@ public class LSBook implements OrderBook {
 	public double returnRate_t = 0;
 
 	// external signal to update threshold
-	public double epsilon_t;
+	public double epsilon_t = 0.5;
 
 	public double pastExcessDemand = 0;
 

### 2008-05-03 src/model/agents/FarmerPatientPlayer.java
Adds support to Farmer model for rounding prices to cents (or other units) and allows either log or linear pricing outputs for comparison. Changes calculation of Cont volume numbers. Adds example main.properties files for farmer and cont models.

@@ -20,8 +20,12 @@ public class FarmerPatientPlayer extends GenericPlayer {
 
 	private int sigma; // order size (constant)
 
-	ArrayList<LimitOrder> myOrders;
+	private double granularity = 0.0; // round to tenths, hundredths, thousandths, etc
+	
+	private boolean useLogPricing = true;
 
+	ArrayList<LimitOrder> myOrders;
+	
 	public FarmerPatientPlayer() {
 		myOrders = new ArrayList<LimitOrder>();
 	}
@@ -35,6 +39,8 @@ public class FarmerPatientPlayer extends GenericPlayer {
 		alpha = myWorld.parameterMap.get("Farmer_alpha");
 		delta = myWorld.parameterMap.get("Farmer_delta");
 		sigma = myWorld.parameterMap.get("Farmer_sigma").intValue();
+		granularity = myWorld.parameterMap.get("Farmer_granularity");
+		useLogPricing = myWorld.optionsMap.get("Farmer_logPricing").equalsIgnoreCase("true");
 
 		target.schedule.scheduleRepeating(0.0, 1, this, 1.0);
 	}
@@ -47,6 +53,14 @@ public class FarmerPatientPlayer extends GenericPlayer {
 
 	}
 
+	private double roundPrice(double price) {
+	  if (granularity == 0.0) {
+		  return price;
+	  } else {
+		  return ((double)((int)(price * granularity + 0.5)))/granularity;
+	  }
+	}
+	
 	private void generateOrders() {
 
 		int numOrdersPlaced = randDist.nextPoisson(alpha);
@@ -65,7 +79,14 @@ public class FarmerPatientPlayer extends GenericPlayer {
 				// Exponential distribution: Log prices at "uniform intensity"
 				double offset = randDist.nextExponential(0.9);
 
-				price = ask - Math.log(offset + 1.0);
+				if (useLogPricing) {
+					price = ask - Math.log(offset+1.0);
+				}
+				else {
+					price = Math.exp(Math.log(ask) - Math.log(offset + 1.0));
+					price = roundPrice(price);
+				}
+
 
 			} else {
 				// Make a sale limit order
@@ -75,10 +96,17 @@ public class FarmerPatientPlayer extends GenericPlayer {
 				// Exponential distribution: Log prices at "uniform intensity"
 				double offset = randDist.nextExponential(0.9);
 
-				price = bid + Math.log(offset + 1.0);
+				if (useLogPricing) {
[…]

### 2008-05-03 src/model/market/books/ContBook.java

@@ -34,7 +34,7 @@ public class ContBook implements OrderBook {
 	public void cleanup() {
 		
 
-		oldVolume = Math.min(currentDemand, currentSupply);
+		oldVolume = Math.max(currentDemand, currentSupply);
 		
 		averageTradePrice = oldVolume*price_t;
 		
@@ -139,14 +139,12 @@ public class ContBook implements OrderBook {
 
 
 	public double getVolume() {
-		// TODO Auto-generated method stub
 		return this.oldVolume;
 	}
 
 
 
 	public double getAverageTradePrice() {
-		// TODO Auto-generated method stub
 		return  this.averageTradePrice;
 	}
 	

### 2008-05-03 src/support/ModelFactory.java

@@ -51,11 +51,13 @@ public class ModelFactory {
 		target.parameterMap.put("Farmer_delta", new Double(properties.getProperty("Farmer_delta", ".01")));
 		target.parameterMap.put("Farmer_mu", new Double(properties.getProperty("Farmer_mu", "0.5")));
 		target.parameterMap.put("Farmer_sigma", new Double(properties.getProperty("Farmer_sigma", "1")));
+		target.parameterMap.put("Farmer_granularity", new Double(properties.getProperty("Farmer_granularity", "0.0")));
 
 		target.parameterMap.put("LS_h", new Double(properties.getProperty("LS_h", "1")));
 		target.parameterMap.put("LS_kappa", new Double(properties.getProperty("LS_kappa", "1")));
 		target.optionsMap.put("LS_herding", properties.getProperty("LS_herding", "true"));
 		target.optionsMap.put("LS_volatility", properties.getProperty("LS_volatility", "true"));
+		target.optionsMap.put("Farmer_logPricing", properties.getProperty("Farmer_logPricing", "true"));
 		
 		target.optionsMap.put("agentConfiguration", properties.getProperty("agentConfiguration", "cont.txt"));
 		target.optionsMap.put("orderBookClass", properties.getProperty("orderBookClass", "ContBook"));

### 2008-05-03 src/model/FinancialModel.java
Adds model which is an amalgam of Farmer and Cont. Impatient agents place market orders according to a signal ala Cont. Patient agents place limit orders ala Farmer and they interact on a DoubleAuction. Example setup files are included.

@@ -69,21 +69,19 @@ public class FinancialModel extends SimState {
 			private static final long serialVersionUID = 6184761986120478954L;
 
 			public void step(SimState state) {
-
-				
 				
 				if (!wrapperActive) {
 					myReporter.finishAll();
 				}
 				
 				state.finish();
-
 			}
+
 		};
 		
 		// run the model maxT times and then finish the execution
 		schedule.scheduleOnce(this.parameterMap.get("maxT"), (int) schedule.MAXIMUM_INTEGER, finalAgent);
-		
+	
 		schedule.scheduleRepeating(myReporter, 10, 1);
 	}
 

### 2008-05-06 documents/WCSS2008Paper/paper.tex
Added sequence diagram to the presentation and paper
Renamed plots from Farmer with a 'Farmer-' prefix. Replaced acf and return histogram plots with representative examples. Updated paper.tex to point to these files.

--- 2008-05-01 documents/WCSS2008Paper/paper.tex
@@
 \centering
-\includegraphics[width=1.0\textwidth]{../graphics/masterClassDiagram.jpg}
+\includegraphics[width=1.0\textwidth]{../graphics/masterClassDiagram.pdf}
 \caption{High-level UML class diagram of the main components and relations in the FinancialMarketModel, including the main attributes of Players and OrderBooks.
@@
 \label{fig:general_uml}
+\end{figure}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=1.0\textwidth]{../graphics/masterSequenceDiagram.pdf}
+\caption{High-level UML sequence diagram of the interactions between main object of the FinancialMarketModel.}
+\label{fig:general_uml}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth,height=0.8\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
+      \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Orderbook shape.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-orderbooks.pdf}}} \quad
+      \subfigure[Trade volumes.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-volumesPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Returns histogram.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnsHistogram.pdf}}} \quad
+      \subfigure[Autocorrelation of returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-acfReturns.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation for default Farmer's parametrization.}
+    \label{fig:sampleDynamics}
+  \end{center}
 \end{figure}

### 2008-05-06 scripts/plotComplex.m
Fixed Cont (after some change, wrong probabilities were used when deciding if to update)

@@ -1,0 +1,23 @@
+filename='C:\\MAStools\\workspace\\FinancialModel\\timeSeries.txt';
+[h, run]=hdrload(filename);
+
+index=6;
+lag=10;
+
+
+data=zeros(length(run)-lag+1,2);
+
+data(:,1)=run(lag:length(run),index);
+data(:,2)=run(1:(length(run)-lag+1),index);
+
+ccolors = vector2colors(run(1:(length(run)-lag+1),2));
+scatter(data(:,1),data(:,2),5,ccolors,'filled'); 
+    
+
+hold on;
+
+for i=2:length(run)
+    
+end
+
+hold off;

### 2008-05-06 src/gui/GUIReporter.java

@@ -37,7 +37,7 @@ public class GUIReporter implements Steppable {
 	
 	public ArrayList<Vector<Double>> volumeMemory = new ArrayList<Vector<Double>>();
 	
-	public int NumViewable = 6000; // also, the #/data points used in
+	public int NumViewable = 10000; // also, the #/data points used in
 
 	// calculating acf
 

### 2008-05-06 src/model/agents/ContPlayer.java

@@ -33,13 +33,14 @@ public class ContPlayer extends GenericPlayer {
 		if (state.schedule.getTime() == 0) {
 
 			this.s_local = myWorld.parameterMap.get("Cont_s");
-			
 			// assign a random threshold; values range from 0 to 1
-			this.threshold = myWorld.random.nextDouble();
+			this.threshold = myWorld.parameterMap.get("D") * myWorld.random.nextDouble();
 		}
 
 		if (state.schedule.getTime() > 0) {
+
 			this.updateThresholds();
+
 		}
 
 		this.generateOrders();
@@ -51,7 +52,7 @@ public class ContPlayer extends GenericPlayer {
 	public void generateOrders() {
 
 		LimitOrder tempOrder;
-		
+
 		double epsilon_t = myWorld.myMarket.getRandomComponentForAsset(0);
 
 		// if epsilon_t is below the negative value of threshold

### 2008-05-06 src/model/market/books/ContBook.java

@@ -75,9 +75,9 @@ public class ContBook implements OrderBook {
 	
 
 	public double priceImpact(double d) {
-		return Math.atan(d / myWorld.parameterMap.get("Cont_lambda"));
+		// return Math.atan(d / myWorld.parameterMap.get("Cont_lambda"));
 		// return Math.atan2(d, myWorld.lambda);
-		// return d / myWorld.lambda;
+		return d / myWorld.parameterMap.get("Cont_lambda");
 	}
 
 	public void setMyWorld(FinancialModel myWorld) {

### 2008-05-06 src/support/Reporter.java

@@ -29,13 +29,13 @@ public class Reporter implements Steppable {
 			
 			
 			String temp = "";
-			temp = temp + "runID" + ";";
-			temp = temp + "time" + ";";
-			temp = temp + "asset" + ";";
-			temp = temp + "closeAskPrice" + ";";
-			temp = temp + "closeBidPrice" + ";";
-			temp = temp + "averageTradePrice" + ";";
-			temp = temp + "volume" + ";";
+			temp = temp + "runID" + "	";
+			temp = temp + "time" + "	";
+			temp = temp + "asset" + "	";
+			temp = temp + "closeAskPrice" + "	";
+			temp = temp + "closeBidPrice" + "	";
+			temp = temp + "averageTradePrice" + "	";
+			temp = temp + "volume" + "	";
 			temp = temp + "return" ;
 			
 
@@ -66,13 +66,14 @@ public class Reporter implements Steppable {
 			for (int asset =0; asset < myModel.parameterMap.get("numAssets"); asset++) {
 			
 			String temp = "";
-			temp = temp + myModel.runID + ";";
-			temp = temp + myModel.schedule.getTime() + ";";
-			temp = temp + myModel.myMarket.getAskPriceForAsset(asset)+ ";";
-			temp = temp + myModel.myMarket.getBidPriceForAsset(asset)+ ";";
-			temp = temp + myModel.myMarket.getAverageTradePriceForAsset(asset) + ";";
-			temp = temp + myModel.myMarket.getVolumeForAsset(asset) + ";";
-			temp = temp + myModel.myMarket.getReturnRateForAsset(asset) + ";";
+			temp = temp + myModel.runID + "	";
+			temp = temp + myModel.schedule.getTime() + "	";
+			temp = temp + asset + "	";
+			temp = temp + myModel.myMarket.getAskPriceForAsset(asset)+ "	";
+			temp = temp + myModel.myMarket.getBidPriceForAsset(asset)+ "	";
+			temp = temp + myModel.myMarket.getAverageTradePriceForAsset(asset) + "	";
+			temp = temp + myModel.myMarket.getVolumeForAsset(asset) + "	";
+			temp = temp + myModel.myMarket.getReturnRateForAsset(asset);
 			
 			outPrices.write(temp);
 			outPrices.newLine();			
@@ -83,7 +84,7 @@ public class Reporter implements Steppable {
 			e.printStackTrace();
 		}
 
-	}
+	}	
 
 	public void finishAll() {
 

### 2008-05-06 documents/WCSS2008Paper/paper.tex
First draft of Cont, Farmer, and Farmer-Cont model descriptions added to paper

@@
 \mailss \\
-\url{http://www.si.edu}}
+\url{http://socialcomplexity.gmu.edu}}
 
@@
 \section{Verification of Correctness}
+TODO: show results side by side with original papers' figures
 
@@
 
+\subsection{Cont Model}
+The Cont model \cite{cont2006} is the simplest implemented.
+All traders (agents) follow the same behavioral rules.
+They are heterogeneous in the sense that they are given independently assigned (subjective) volatility thresholds $\theta_i(t)$.
+Each period, all agents recieve a common signal, which can be interpreted as public information or "news", in the form of IID Gaussian random variables $\epsilon_t$.
+Each agent $i$ responds to this signal by selling if $\epsilon_t < -\theta_i(t)$, buying if $\epsilon_t > \theta_i(t)$, and otherwise sitting out the period.
+The market then determines the excess demand and arrives at the market clearing price by means of a market impact function.
+Lastly, with probability $s$, agents update their threshold to match the absolute value of the return rate for the current period.
+Somewhat surprisingly, even this very simple "zero intelligence" model of agent behavior yields fairly realistic market movements.
+TODO Show results.
+TODO Note:cyclicality.
+
+\subsection{Farmer Model}
+The Farmer model \cite{farmer2003} has two types of agents, which trade a single asset by means of a continuous double auction similar to those normally used on stock markets.
+"Patient" traders place orders place limit orders, which specify both a price and quantity they wish to buy or sell.
+Limit orders do not execute immediately, and may be canceled at any time.
+"Impatient" traders place market orders, specifying only a quantity, which execute immediately at the current best price, assuming there is sufficient liquidity provided by outstanding limit orders.
+This model also has simple behavioral logic for traders.
+Behavior is regulated by:
 \begin{itemize}
-	\item \cite{farmer2003}
-	\item \cite{lamba2007}
-	\item \cite{westerhoff2004}
-	\item \cite{cont2006}
+\item Patient agents place limit orders of constant size at a Poisson rate per unit time, and with a randomly generated price.
+Orders are equally likely to be either buy or sell orders.
+Buy limit orders are placed uniformly anywhere in the semiinfinite interval below the current ask price ($-\infty < p < a(t)$), where $p$ is the \emph{logarithm} of the price; similarly for sell orders.
+A uniform distribution in log prices is equivalent to an exponential distribution in prices.
+In addition, outstanding limit orders are cancelled at a Poisson rate per unit time.
+All of these processes are independent.
+\item Impatient agents place market orders of constant size at a Poisson rate per unit time.
+\item A double auction order book.
+This manages the pending limit orders and facilitates trades between market and limit orders.
+It also provides aggregate market information such as return rate per unit time and bid-ask spreads.
 \end{itemize}
+This model explains a large portion of the spread and price diffusion of actual stocks given an order flow rate, which essentially means that the double auction structure has a large impact on the nature of market movements.
+The model also yields market behavior qualitatively close to actual markets in many respects.
+TODO Show results. 
+TODO Note: log vs. non-log prices
+TODO Note: effect of price granularity
+
+\subsection{Farmer-Cont Model}
+We constructed a model combining aspects of the Farmer model with aspects of the Cont model.
+The result is a model which is still essentially zero intelligence, but which exhibits empirically positive characteristics of both models.
+In this model there are two types of agents, which interact via the continuous double auction mechanism, as in Farmer.
+In fact, the Patient traders, which place limit orders and thus provide liquidity, remain the same.
+Impatient traders are modeled after Cont.
+They have heterogeneous volatility thresholds, and place market orders depending upon how public information in the form of a normally distributed random variable compares with these thresholds.
+TODO Show results.
+TODO Note: liquidity issues
+
+\subsection{Lambda Model}
+\cite{lamba2007}
+TODO ...
+
+\subsection{Westerhoff Model}
+\cite{westerhoff2004}
+TODO ...
 

### 2008-05-06 src/gui/FinancialModelWithUI.java
1) Small bug updates; 2) Set background of all charts to white;

@@ -7,6 +7,8 @@
 package gui;
 
 import java.awt.BorderLayout;
+import java.awt.Color;
+import java.awt.Paint;
 
 import javax.swing.JFrame;
 
@@ -27,7 +29,7 @@ public class FinancialModelWithUI extends GUIState {
 	public TimeSeriesChartGenerator priceChart;
 
 	public JFrame priceFrame;
-	
+
 	public TimeSeriesChartGenerator returnChart;
 
 	public JFrame returnFrame;
@@ -35,7 +37,7 @@ public class FinancialModelWithUI extends GUIState {
 	public TimeSeriesChartGenerator acfChart;
 
 	public JFrame acfFrame;
-	
+
 	public TimeSeriesChartGenerator volumeChart;
 
 	public JFrame volumeFrame;
@@ -99,7 +101,7 @@ public class FinancialModelWithUI extends GUIState {
 
 					if (returnHistFrame.isVisible()) {
 						for (int a = 0; a < ((FinancialModel) state).parameterMap.get("numAssets"); a++) {
-							
+
 							Double[] tempArray = new Double[myReporter.returnMemory.get(a).size()];
 							tempArray = myReporter.returnMemory.get(a).toArray(tempArray);
 							double[] temp2Array = new double[myReporter.returnMemory.get(a).size()];
@@ -114,17 +116,22 @@ public class FinancialModelWithUI extends GUIState {
 
 					if (orderHistFrame.isVisible()) {
 						for (int a = 0; a < ((FinancialModel) state).parameterMap.get("numAssets"); a++) {
-							
-							//((FinancialModel) state).myMarket.orderBooks.get(a).cleanup();
-							
+
+							// ((FinancialModel)
+							// state).myMarket.orderBooks.get(a).cleanup();
+
 							double[] temp2Array = ((FinancialModel) state).myMarket.orderBooks.get(a).getBuyOrders();
-							if (temp2Array.length > 0) {
-								orderHist.updateSeries(2 * a, temp2Array, false);
+							if (temp2Array != null) {
+								if (temp2Array.length > 0) {
+									orderHist.updateSeries(2 * a, temp2Array, false);
+								}
 							}
 
 							temp2Array = ((FinancialModel) state).myMarket.orderBooks.get(a).getSellOrders();
-							if (temp2Array.length > 0) {
-								orderHist.updateSeries(2 * a + 1, temp2Array, false);
[…]

### 2008-05-06 src/gui/GUIReporter.java

@@ -37,7 +37,7 @@ public class GUIReporter implements Steppable {
 	
 	public ArrayList<Vector<Double>> volumeMemory = new ArrayList<Vector<Double>>();
 	
-	public int NumViewable = 10000; // also, the #/data points used in
+	public int NumViewable = 2000; // also, the #/data points used in
 
 	// calculating acf
 

### 2008-05-06 documents/WCSS2008Paper/paper.tex
Added graphs for Cont's default dynamics

@@
 
-\begin{figure}[htbp]
-  \begin{center}
-   \mbox{
-      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth,height=0.8\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
-      \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnPlot.pdf}}}
-      }
-    \mbox{
-      \subfigure[Orderbook shape.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-orderbooks.pdf}}} \quad
-      \subfigure[Trade volumes.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-volumesPlot.pdf}}}
-      }
-    \mbox{
-      \subfigure[Returns histogram.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnsHistogram.pdf}}} \quad
-      \subfigure[Autocorrelation of returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-acfReturns.pdf}}}
-      }
-    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation for default Farmer's parametrization.}
-    \label{fig:sampleDynamics}
-  \end{center}
-\end{figure}
-
-\section{Verification of Correctness}
-TODO: show results side by side with original papers' figures
-
 \section{Overview of Implemented Models}
@@
 \subsection{Cont Model}
+
 The Cont model \cite{cont2006} is the simplest implemented.
@@
 Lastly, with probability $s$, agents update their threshold to match the absolute value of the return rate for the current period.
-Somewhat surprisingly, even this very simple "zero intelligence" model of agent behavior yields fairly realistic market movements.
-TODO Show results.
+Somewhat surprisingly, even this very simple "zero intelligence" model of agent behavior yields fairly realistic market movements, as presented on Figure \ref{fig:sampleDynamicsCont}.
+
 TODO Note:cyclicality.
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth]{../graphics/Cont-pricesPlot.pdf}}} \quad
+      \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-returnsPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Trade volumes.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-volumesPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Returns histogram.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-returnsHistogram.pdf}}} \quad
+      \subfigure[Autocorrelation of returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-acfPlot.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation for default Cont's parametrization.}
+    \label{fig:sampleDynamicsCont}
+  \end{center}
+\end{figure}
 
@@
 This model explains a large portion of the spread and price diffusion of actual stocks given an order flow rate, which essentially means that the double auction structure has a large impact on the nature of market movements.
-The model also yields market behavior qualitatively close to actual markets in many respects.
-TODO Show results. 
+The model also yields market behavior qualitatively close to actual markets in many respects, see Figure \ref{fig:sampleDynamicsFarmer}.
+
 TODO Note: log vs. non-log prices
 TODO Note: effect of price granularity
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
+      \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Orderbook shape.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-orderbooks.pdf}}} \quad
+      \subfigure[Trade volumes.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-volumesPlot.pdf}}}
+      }
+    \mbox{
+      \subfigure[Returns histogram.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnsHistogram.pdf}}} \quad
+      \subfigure[Autocorrelation of returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-acfReturns.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation for default Farmer's parametrization.}
+    \label{fig:sampleDynamicsFarmer}
+  \end{center}
+\end{figure}
 
@@
 
-\subsection{Lambda Model}
-\cite{lamba2007}
-TODO ...
-
-\subsection{Westerhoff Model}
-\cite{westerhoff2004}
-TODO ...
-
+\section{Verification of Correctness}
+TODO: show results side by side with original papers' figures
 \bibliographystyle{plainnat}

### 2008-05-07 documents/WCSS2008Paper/paper.tex
Adds graphics for two parameter values of Cont model, comparisons with Cont's published results, and an explanation of periodicity in the paper.

--- 2008-05-06 documents/WCSS2008Paper/paper.tex
@@
 
+\section{Overview of Implemented Models}
+
+\subsection{Cont Model}
+
+The Cont model (\cite{cont2006}) is the simplest implemented.
+All traders (agents) follow the same behavioral rules.
+They are heterogeneous in the sense that they are given independently assigned (subjective) volatility thresholds $\theta_i(t)$.
+Each period, all agents recieve a common signal, which can be interpreted as public information or "news", in the form of IID Gaussian random variables $\epsilon_t ~ N(0,D^2)$.
+Each agent $i$ responds to this signal by selling if $\epsilon_t < -\theta_i(t)$, buying if $\epsilon_t > \theta_i(t)$, and otherwise sitting out the period.
+The market then determines the excess demand and arrives at the market clearing price by means of a market impact function.
+Lastly, with probability $s$, agents update their threshold to match the absolute value of the return rate for the current period.
+Somewhat surprisingly, even this very simple "zero intelligence" model of agent behavior yields fairly realistic market movements, as presented in Figures \ref{fig:sampleDynamicsCont_s01} and \ref{fig:sampleDynamicsCont_s1}.
+These compare well with data shown in the original paper (shown in Figures \ref{fig:ContFig3} and \ref{fig:ContFig4}).
+
+One down side of the Cont model that is not noted in the papers is that there appears to be a great degree of periodic cyclicality in the volatility of the markets.
+This is especially apparent with higher probability $s$ of updating thresholds.
+With $s=.1$, this is readily apparent if you look at the autocorrelation of absolute returns over longer time scales than shown in the paper (Figure \ref{fig:ContACF1}).
+This effect can also be seen when averaging over many runs (See Figure \ref{fig:ContACF2}).
+Holding $s$ constant at 1 and varying $D$, we find the period and amplitude to be predictable.
+
 \begin{figure}[htbp]
@@
    \mbox{
-      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth,height=0.8\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
+    \subfigure[Trade volumes.]{\scalebox{0.33}{ \includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-volumesPlot.pdf}}}
+    \quad
+\subfigure[Price]{\scalebox{0.33}{\includegraphics[width=1.0\textwidth]{../graphics/Cont-s01-pricesPlot.pdf}}}
+}
+    \mbox{
+      \subfigure[Raw and absolute log returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsPlot.pdf}}}
+      \quad
+      \subfigure[Distribution of Returns]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsHistogram.pdf}}}       
+      \quad
+      \subfigure[Distribution of Returns (log scale)]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-logReturnsHistogram.pdf}}}       
+}
+    \mbox{
+      \subfigure[Annualized moving average volatility]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-avgVolatility.pdf}}}
+      \quad
+      \subfigure[Autocorrelation of returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-acfPlot-noAbs.pdf}}}
+      \quad
+      \subfigure[Autocorrelation of absolute returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-acfPlot-absonly.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the Cont FinancialModel simulation with $D=0.001$, $s=0.01$.}
+    \label{fig:sampleDynamicsCont_s01}
+  \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=1.00\textwidth]{../graphics/Cont-Fig3.png}
+    \caption{Corresponding figure from Cont's simulation with $D=0.001$, $s=0.01$.}
+    \label{fig:ContFig3}
+  \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Trade volumes.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-volumesPlot.pdf}}}
+      \quad
+      \subfigure[Price]{\scalebox{0.33}{ \includegraphics[width=1.0\textwidth]{../graphics/Cont-pricesPlot.pdf}}} \quad
+      }
+    \mbox{
+      \subfigure[Raw and absolute log returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-returnsPlot.pdf}}}
+      \quad
+      \subfigure[Returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-returnsHistogram.pdf}}}
+      \quad
+      \subfigure[Distribution of Returns (log scale)]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-logReturnsHistogram.pdf}}}       
+}
+    \mbox{
+      \subfigure[Annualized moving average volatility]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-avgVolatility.pdf}}}
+      \quad
+      \subfigure[Autocorrelation of returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-acfPlot-noAbs.pdf}}}
+      \quad
+      \subfigure[Autocorrelation of absolute returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-acfPlot-absonly-to50.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the Cont FinancialModel simulation with $D=0.001$, $s=0.1$.}
+    \label{fig:sampleDynamicsCont_s1}
+  \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=1.0\textwidth]{../graphics/Cont-Fig4.png}
+    \caption{Corresponding figure from Cont's simulation with $D=0.001$, $s=0.1$.}
+    \label{fig:ContFig4}
+  \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+   \includegraphics[width=0.9\textwidth]{../graphics/Cont-acfPlot.pdf}
+    \caption{Autocorrelation in Cont simulation over a longer time scale, with $D=0.001$, $s=0.1$.}
+    \label{fig:ContACF1}
+  \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}   \includegraphics[width=1.00\textwidth]{../graphics/comparisonOfACFs.pdf}
+    \caption{Autocorrelation in Cont simulation $s=0.1$ and varying $D$, averaged over many runs.}
+    \label{fig:ContACF2}
+  \end{center}
+\end{figure}
+
+\subsection{Farmer Model}
+The Farmer model (\cite{farmer2003}) has two types of agents, which trade a single asset by means of a continuous double auction similar to those normally used on stock markets.
+"Patient" traders place limit orders, which specify both a price and quantity they wish to buy or sell.
+Limit orders do not execute immediately, and may be canceled at any time.
+"Impatient" traders place market orders, specifying only a quantity, which execute immediately at the current best price, assuming there is sufficient liquidity provided by outstanding limit orders.
+This model also has simple behavioral logic for traders.
+Behavior is regulated by:
+\begin{itemize}
+\item Patient agents placing limit orders of constant size at a Poisson rate per unit time, and with a randomly generated price.
+Orders are equally likely to be either buy or sell orders.
+Buy limit orders are placed uniformly anywhere in the semiinfinite interval below the current ask price ($-\infty < p < a(t)$), where $p$ is the \emph{logarithm} of the price; similarly for sell orders.
+A uniform distribution in log prices is equivalent to an exponential distribution in prices.
+In addition, outstanding limit orders are cancelled at a Poisson rate per unit time.
+All of these processes are independent.
+\item Impatient agents placing market orders of constant size at a Poisson rate per unit time.
+\item A double auction order book.
+This manages the pending limit orders and facilitates trades between market and limit orders.
+It also provides aggregate market information such as return rate per unit time and bid-ask spreads.
+\end{itemize}
+This model explains a large portion of the spread and price diffusion of actual stocks given an order flow rate, which essentially means that the double auction structure has a large impact on the nature of market movements.
+The model also yields market behavior qualitatively close to actual markets in many respects, see Figure \ref{fig:sampleDynamicsFarmer}.
+
+TODO Note: log vs. non-log prices
+TODO Note: effect of price granularity
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
       \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnPlot.pdf}}}
@@
     \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation for default Farmer's parametrization.}
-    \label{fig:sampleDynamics}
+    \label{fig:sampleDynamicsFarmer}
   \end{center}
@@
 
-\section{Verification of Correctness}
-TODO: show results side by side with original papers' figures
-
-\section{Overview of Implemented Models}
-
-\subsection{Cont Model}
-The Cont model \cite{cont2006} is the simplest implemented.
-All traders (agents) follow the same behavioral rules.
-They are heterogeneous in the sense that they are given independently assigned (subjective) volatility thresholds $\theta_i(t)$.
-Each period, all agents recieve a common signal, which can be interpreted as public information or "news", in the form of IID Gaussian random variables $\epsilon_t$.
-Each agent $i$ responds to this signal by selling if $\epsilon_t < -\theta_i(t)$, buying if $\epsilon_t > \theta_i(t)$, and otherwise sitting out the period.
-The market then determines the excess demand and arrives at the market clearing price by means of a market impact function.
-Lastly, with probability $s$, agents update their threshold to match the absolute value of the return rate for the current period.
-Somewhat surprisingly, even this very simple "zero intelligence" model of agent behavior yields fairly realistic market movements.
-TODO Show results.
-TODO Note:cyclicality.
-
-\subsection{Farmer Model}
-The Farmer model \cite{farmer2003} has two types of agents, which trade a single asset by means of a continuous double auction similar to those normally used on stock markets.
-"Patient" traders place orders place limit orders, which specify both a price and quantity they wish to buy or sell.
-Limit orders do not execute immediately, and may be canceled at any time.
-"Impatient" traders place market orders, specifying only a quantity, which execute immediately at the current best price, assuming there is sufficient liquidity provided by outstanding limit orders.
-This model also has simple behavioral logic for traders.
-Behavior is regulated by:
-\begin{itemize}
-\item Patient agents place limit orders of constant size at a Poisson rate per unit time, and with a randomly generated price.
-Orders are equally likely to be either buy or sell orders.
-Buy limit orders are placed uniformly anywhere in the semiinfinite interval below the current ask price ($-\infty < p < a(t)$), where $p$ is the \emph{logarithm} of the price; similarly for sell orders.
-A uniform distribution in log prices is equivalent to an exponential distribution in prices.
-In addition, outstanding limit orders are cancelled at a Poisson rate per unit time.
-All of these processes are independent.
-\item Impatient agents place market orders of constant size at a Poisson rate per unit time.
-\item A double auction order book.
-This manages the pending limit orders and facilitates trades between market and limit orders.
-It also provides aggregate market information such as return rate per unit time and bid-ask spreads.
-\end{itemize}
-This model explains a large portion of the spread and price diffusion of actual stocks given an order flow rate, which essentially means that the double auction structure has a large impact on the nature of market movements.
-The model also yields market behavior qualitatively close to actual markets in many respects.
-TODO Show results. 
-TODO Note: log vs. non-log prices
-TODO Note: effect of price granularity
-
 \subsection{Farmer-Cont Model}
-We constructed a model combining aspects of the Farmer model with aspects of the Cont model.
+As an experiment, we constructed a model combining some aspects of the Farmer model with some aspects of the Cont model.
 The result is a model which is still essentially zero intelligence, but which exhibits empirically positive characteristics of both models.
@@
 They have heterogeneous volatility thresholds, and place market orders depending upon how public information in the form of a normally distributed random variable compares with these thresholds.
+Some fraction $s$ of agents then adjust their volatility thresholds in light of current volatility by setting it to the trailing period's return rate.
 TODO Show results.

### 2008-05-07 src/gui/FinancialModelWithUI.java
Computes and adds a new graph of trailing average annual volatility (absolute returns). This is another of the statistics shown by Cont. Reverts price impact to use atan, since it qualitatively matches the results Cont published, better than the other example he gave.

@@ -34,6 +34,10 @@ public class FinancialModelWithUI extends GUIState {
 
 	public JFrame returnFrame;
 
+	public TimeSeriesChartGenerator trailingAbsReturnChart;
+
+	public JFrame trailingReturnFrame;
+	
 	public TimeSeriesChartGenerator acfChart;
 
 	public JFrame acfFrame;
@@ -139,7 +143,12 @@ public class FinancialModelWithUI extends GUIState {
 					if (acfFrame.isVisible()) {
 						myReporter.acfAbsReturns();
 					}
+					
+					if (trailingReturnFrame.isVisible()) {
+						myReporter.updateTrailing();
+					}
 
+					// TODO: why set series three times??
 					if (priceFrame.isVisible()) {
 						priceChart.disable();
 						myReporter.setSeries();
@@ -203,6 +212,21 @@ public class FinancialModelWithUI extends GUIState {
 		returnFrame.pack();
 		c.registerFrame(returnFrame);
 
+		trailingAbsReturnChart = new TimeSeriesChartGenerator();
+		trailingAbsReturnChart.setTitle("Annual average volatility plot");
+		trailingAbsReturnChart.setDomainAxisLabel("Step");
+		trailingAbsReturnChart.setRangeAxisLabel("Trailing average annual absolute returns");
+		trailingAbsReturnChart.getChartPanel().getChart().setBackgroundPaint(Color.white);
+		for (int a = 0; a < myModel.parameterMap.get("numAssets"); a++) {
+			trailingAbsReturnChart.addSeries(myReporter.trailingAbsReturnSeries.get(a), null);
+		}
+		trailingReturnFrame = trailingAbsReturnChart.createFrame(this);
+		trailingReturnFrame.getContentPane().setLayout(new BorderLayout());
+		trailingReturnFrame.getContentPane().add(trailingAbsReturnChart, BorderLayout.CENTER);
+		trailingReturnFrame.pack();
+		c.registerFrame(trailingReturnFrame);
+		
+		
 		volumeChart = new TimeSeriesChartGenerator();
 		volumeChart.setTitle("Volumes plot");
 		volumeChart.setDomainAxisLabel("Step");
@@ -238,7 +262,7 @@ public class FinancialModelWithUI extends GUIState {
 		returnHist.setTitle("Returns histogram");
 		returnHist.setDomainAxisLabel("Value");
 		returnHist.setRangeAxisLabel("Number of observations");
[…]

### 2008-05-07 src/gui/GUIReporter.java

@@ -23,6 +23,8 @@ public class GUIReporter implements Steppable {
 
 	public ArrayList<XYSeries> absReturnSeries = new ArrayList<XYSeries>();
 
+	public ArrayList<XYSeries> trailingAbsReturnSeries = new ArrayList<XYSeries>();
+	
 	public ArrayList<XYSeries> acfAbsReturnsSeries = new ArrayList<XYSeries>();
 
 	public ArrayList<XYSeries> acfReturnsSeries = new ArrayList<XYSeries>();
@@ -37,12 +39,14 @@ public class GUIReporter implements Steppable {
 	
 	public ArrayList<Vector<Double>> volumeMemory = new ArrayList<Vector<Double>>();
 	
-	public int NumViewable = 2000; // also, the #/data points used in
+	public int NumViewable = 10000; // also, the #/data points used in
 
 	// calculating acf
 
 	public int ACFViewable = 200;
 
+	public int trailingSampleSize = 365;
+	
 	FinancialModel myModel;
 
 	public int nextUpdate = 0;
@@ -57,6 +61,7 @@ public class GUIReporter implements Steppable {
 			bidPriceSeries.add(new XYSeries("Bid price for asset " + a));
 			askPriceSeries.add(new XYSeries("Ask price for asset " + a));
 			returnSeries.add(new XYSeries("Returns for asset " + a));
+			trailingAbsReturnSeries.add(new XYSeries("Trailing volatility for asset " + a));
 			absReturnSeries.add(new XYSeries("Absolute returns for asset " + a));
 			volumeSeries.add(new XYSeries("Trade volumes for asset " + a));
 			acfAbsReturnsSeries.add(new XYSeries("ACF of absolute returns for asset " + a));
@@ -66,12 +71,16 @@ public class GUIReporter implements Steppable {
 			bidPriceSeries.get(a).setMaximumItemCount(NumViewable);
 			askPriceSeries.get(a).setMaximumItemCount(NumViewable);
 			returnSeries.get(a).setMaximumItemCount(NumViewable);
+			trailingAbsReturnSeries.get(a).setMaximumItemCount(NumViewable);
 			volumeSeries.get(a).setMaximumItemCount(NumViewable);
 			absReturnSeries.get(a).setMaximumItemCount(NumViewable);
 			// initialize
 			for (int i = 1; i < ACFViewable; i++) {
[…]

### 2008-05-07 src/model/market/books/ContBook.java

@@ -75,9 +75,9 @@ public class ContBook implements OrderBook {
 	
 
 	public double priceImpact(double d) {
-		// return Math.atan(d / myWorld.parameterMap.get("Cont_lambda"));
+		return Math.atan(d / myWorld.parameterMap.get("Cont_lambda"));
 		// return Math.atan2(d, myWorld.lambda);
-		return d / myWorld.parameterMap.get("Cont_lambda");
+		//return d / myWorld.parameterMap.get("Cont_lambda");
 	}
 
 	public void setMyWorld(FinancialModel myWorld) {

### 2008-05-07 documents/WCSS2008Paper/paper.tex
Modified the paper, locking it for next 2h.
Current state of the paper (worked on the first half). Going for coffee, keeping locks.

@@
 
-\author{Robert Axtell$^{a}$ \and CSS 739 Class Project Team}
+\author{Robert Axtell$^{a}$ \and CSS 739 Class Project Team\thanks{Programming Group: Justin Briggs, Francesco Trusiano, Maciej \L atek, Randy Marks.
+Data Analysis Group: Sandler Lin, Cheryl Hansen and Thomas Tzitzouris.}}
 
@@
 \maketitle
-\begin{abstract}
-AAA
-\end{abstract}
 
@@
 
+The purpose of this project is to create computational platform that would be used to implement agent-based models of financial markets. In particular, following functionalities are desired:
+
+\begin{itemize}
+	\item Simulation may host heterogeneous agents that follow news processes, learn, obey budget constraints etc.;
+	\item Many different assets traded on markets using different trading rules may be present (based on price-impact heuristics, double auction);
+	\item Flexible GUI and reporting facilities, including ability to summarize batch results from sensitivity analysis.
+\end{itemize}
+
+This paper outlines the prototype prepared during the CSS739 class, presents three sample models and follows the steps taken to ensure correctness of our implementation.
+
 \section{Platform Architecture}
+
+The implementation took place using MASON simulation framework, see \cite{luke2005}.
+The static structure of our platform is outlined on Figure \ref{fig:general_class}, where UML class diagram is presented (minus classes corresponding to orders).
+Figure \ref{fig:general_sequence} presents interactions taking place during a sample iteration between two agents and an orderbook.
 
@@
 Agent classes (light blue) inherit from the MASON \texttt{Steppable} interface while the master class is implementation of MASON's \texttt{SimState}.}
-\label{fig:general_uml}
+\label{fig:general_class}
 \end{figure}
@@
 \includegraphics[width=1.0\textwidth]{../graphics/masterSequenceDiagram.pdf}
-\caption{High-level UML sequence diagram of the interactions between main object of the FinancialMarketModel.}
-\label{fig:general_uml}
+\caption{High-level UML sequence diagram of the interactions between main object of the FinancialMarketModel, with messages send between two agents and a single orderbook during a single tick shown.
+Agents are activated by MASON simulation engine and may generate different kinds of \texttt{Market}- and \texttt{LimitOrders}, depending on their internal states and signals from environment (prices, news process).
+The \texttt{MarketOrder} gives an agent instantaneous feedback, whereupon effects of LimitOrder may be delayed in time.
+The \texttt{cleanup} signal causes the Orderbook to arbitrage all possible trades between \texttt{LimitOrders}, recalculate new prices and returns (if applicable) and remove expired orders.}
+\label{fig:general_sequence}
 \end{figure}
 
-\section{Overview of Implemented Models}
+\section{Implemented Models}
 
@@
 
-The Cont model (\cite{cont2006}) is the simplest implemented.
+The Cont model (\cite{cont2006}) is the simplest model implemented.
 All traders (agents) follow the same behavioral rules.
 They are heterogeneous in the sense that they are given independently assigned (subjective) volatility thresholds $\theta_i(t)$.
-Each period, all agents recieve a common signal, which can be interpreted as public information or "news", in the form of IID Gaussian random variables $\epsilon_t ~ N(0,D^2)$.
+Each period, all agents receive a common signal, which can be interpreted as public information or "news", in the form of IID Gaussian random variables $\epsilon_t ~ N(0,D^2)$.
 Each agent $i$ responds to this signal by selling if $\epsilon_t < -\theta_i(t)$, buying if $\epsilon_t > \theta_i(t)$, and otherwise sitting out the period.
@@
 Lastly, with probability $s$, agents update their threshold to match the absolute value of the return rate for the current period.
-Somewhat surprisingly, even this very simple "zero intelligence" model of agent behavior yields fairly realistic market movements, as presented in Figures \ref{fig:sampleDynamicsCont_s01} and \ref{fig:sampleDynamicsCont_s1}.
-These compare well with data shown in the original paper (shown in Figures \ref{fig:ContFig3} and \ref{fig:ContFig4}).
+Somewhat surprisingly, even this very simple "zero intelligence" model of agent behavior yields fairly realistic market movements, as presented in Figures \ref{fig:ContSmallSSim} (for small $s$) and \ref{fig:ContLargeSSim} (for large value of $s$).
 
@@
 This is especially apparent with higher probability $s$ of updating thresholds.
-With $s=.1$, this is readily apparent if you look at the autocorrelation of absolute returns over longer time scales than shown in the paper (Figure \ref{fig:ContACF1}).
-This effect can also be seen when averaging over many runs (See Figure \ref{fig:ContACF2}).
-Holding $s$ constant at 1 and varying $D$, we find the period and amplitude to be predictable.
+With $s=0.1$, this is readily apparent if you look at the autocorrelation of absolute returns over longer time scales than shown in the paper.
+This effect can be particularly well seen when averaging over many runs, as on Figure \ref{fig:ContMultiRun}.
+Holding $s$ constant at $0.1$ and varying $D$, we find the period and amplitude to be predictable.
 
@@
     \caption{Examples of outputs and statistics from a single run of the Cont FinancialModel simulation with $D=0.001$, $s=0.01$.}
-    \label{fig:sampleDynamicsCont_s01}
-  \end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=1.00\textwidth]{../graphics/Cont-Fig3.png}
-    \caption{Corresponding figure from Cont's simulation with $D=0.001$, $s=0.01$.}
-    \label{fig:ContFig3}
+    \label{fig:ContSmallSSim}
   \end{center}
@@
     \caption{Examples of outputs and statistics from a single run of the Cont FinancialModel simulation with $D=0.001$, $s=0.1$.}
-    \label{fig:sampleDynamicsCont_s1}
-  \end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width=1.0\textwidth]{../graphics/Cont-Fig4.png}
-    \caption{Corresponding figure from Cont's simulation with $D=0.001$, $s=0.1$.}
-    \label{fig:ContFig4}
-  \end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
-   \includegraphics[width=0.9\textwidth]{../graphics/Cont-acfPlot.pdf}
-    \caption{Autocorrelation in Cont simulation over a longer time scale, with $D=0.001$, $s=0.1$.}
-    \label{fig:ContACF1}
-  \end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}   \includegraphics[width=1.00\textwidth]{../graphics/comparisonOfACFs.pdf}
-    \caption{Autocorrelation in Cont simulation $s=0.1$ and varying $D$, averaged over many runs.}
-    \label{fig:ContACF2}
+    \label{fig:ContLargeSSim}
   \end{center}
@@
 \subsection{Farmer Model}
+
 The Farmer model (\cite{farmer2003}) has two types of agents, which trade a single asset by means of a continuous double auction similar to those normally used on stock markets.
@@
 \subsection{Farmer-Cont Model}
+
 As an experiment, we constructed a model combining some aspects of the Farmer model with some aspects of the Cont model.
@@
 \section{Verification of Correctness}
+
 TODO: show results side by side with original papers' figures
+
+These compare well with data shown in the original paper (shown in Figures \ref{fig:ContSmallSPap} and \ref{fig:ContLargeSPap}).
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{../graphics/Cont-Fig3.png}
+    \caption{Outputs from Cont's paper with $D=0.001$, $s=0.01$. Compare directly to Figure \ref{fig:ContSmallSSim}.}
+    \label{fig:ContSmallSPap}
+  \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{../graphics/Cont-Fig4.png}
+    \caption{Outputs from Cont's paper with $D=0.001$, $s=0.1$. Compare directly to Figure \ref{fig:ContLargeSSim}.}
+    \label{fig:ContLargeSPap}
+  \end{center}
+\end{figure}
+
+\section{Muti-run Instrumentation}
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Autocorrelation functions]{\scalebox{0.5}{\quad \quad \includegraphics[width=1.0\textwidth]{../graphics/comparisonOfACFs.pdf}}} \quad
+      \subfigure[Return distributions]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/comparisonOfHistograms.pdf}}}
+      }
+    \caption{Autocorrelation and distributions in Cont simulation with $s=0.1$ and varying $D$, averaged over many runs.}
+    \label{fig:ContMultiRun}
+  \end{center}
+\end{figure}
+
 \bibliographystyle{plainnat}

### 2008-05-07 documents/WCSS2008Paper/paper.tex
1) Corrected histograms for all Cont graphs 2) Added diagnostics for FC
Keeping lock on tex file

@@
     \mbox{
-      \subfigure[Raw and absolute log returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsPlot.pdf}}}
+      \subfigure[Raw and absolute returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsPlot.pdf}}}
       \quad
-      \subfigure[Distribution of Returns]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsHistogram.pdf}}}       
+      \subfigure[Returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsHistogram.pdf}}}       
       \quad
-      \subfigure[Distribution of Returns (log scale)]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-logReturnsHistogram.pdf}}}       
+      \subfigure[Log returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-logReturnsHistogram.pdf}}}       
 }
@@
       \quad
-      \subfigure[Distribution of Returns (log scale)]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-logReturnsHistogram.pdf}}}       
+      \subfigure[Log returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-logReturnsHistogram.pdf}}}       
 }
@@
 \begin{itemize}
-\item Patient agents placing limit orders of constant size at a Poisson rate per unit time, and with a randomly generated price.
+\item Patient agents place \texttt{LimitOrders} of constant size at a Poisson rate per unit time, and with a randomly generated price.
 Orders are equally likely to be either buy or sell orders.
@@
 A uniform distribution in log prices is equivalent to an exponential distribution in prices.
-In addition, outstanding limit orders are cancelled at a Poisson rate per unit time.
+In addition, outstanding limit orders are canceled at a Poisson rate per unit time.
 All of these processes are independent.
-\item Impatient agents placing market orders of constant size at a Poisson rate per unit time.
+\item Impatient agents place \texttt{MarketOrders} of constant size at a certain Poisson rate per unit time.
 \item A double auction order book.
@@
 
-TODO Note: log vs. non-log prices
-TODO Note: effect of price granularity
-
 \begin{figure}[htbp]
@@
    \mbox{
-      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=0.9\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
+      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=1.0\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
       \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnPlot.pdf}}}
@@
 Some fraction $s$ of agents then adjust their volatility thresholds in light of current volatility by setting it to the trailing period's return rate.
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Trade volumes.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-volumesPlot.pdf}}}
+      \quad
+      \subfigure[Price]{\scalebox{0.33}{ \includegraphics[width=1.0\textwidth]{../graphics/FC-pricesPlot.pdf}}} \quad
+      }
+    \mbox{
+      \subfigure[Raw and absolute log returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-returnsPlot.pdf}}}
+      \quad
+      \subfigure[Returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-returnsHistogram.pdf}}}
+      \quad
+      \subfigure[Returns histogram in logs.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-logReturnsHistogram.pdf}}}       
+}
+    \mbox{
+      \subfigure[Annualized moving average volatility]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-avgVolatility.pdf}}}
+      \quad
+      \subfigure[Snapshot of orderbook.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-orderbooks.pdf}}}
+      \quad
+      \subfigure[Autocorrelation of absolute returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-acfReturns.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the Farmer-Cont hybryd.}
+    \label{fig:ContLargeSSim}
+  \end{center}
+\end{figure}
+
 TODO Show results.

### 2008-05-07 documents/WCSS2008Paper/paper.tex

@@
       }
-    \caption{Examples of outputs and statistics from a single run of the Farmer-Cont hybryd.}
+    \caption{Examples of outputs and statistics from a single run of the Farmer-Cont hybrid.}
     \label{fig:ContLargeSSim}
@@
 
-TODO Show results.
 TODO Note: liquidity issues
@@
 
-\section{Muti-run Instrumentation}
+\section{Batch Mode Instrumentation}
+
+FinancialMarketModel includes a set of wrappers that are useful for conducting multi-run batch parameter space exploration or sensitivity analysis.
+A simple wrapper is included in \texttt{experiments} package.
+Post-run analysis can be performed by using Matlab, sample scripts are included in scripts folder and can be used to render shape of autocorrelation curves and histograms of returns as a function of simulation settings.
+Figure \ref{fig:ContMultiRun} presents a sample prepared by varying the $D$ news process variance for Cont's model.
 

### 2008-05-08 scripts/plotComplex.m
Added some Matlab scripts useful for sensitivity analysis

@@ -1,23 +1,51 @@
-filename='C:\\MAStools\\workspace\\FinancialModel\\timeSeries.txt';
-[h, run]=hdrload(filename);
+filename='..\\timeSeries.txt';
+[h, allRuns]=hdrload(filename);
 
-index=6;
-lag=10;
+maxRuns=100;
 
+hursts=zeros(maxRuns,1);
 
-data=zeros(length(run)-lag+1,2);
+hold all;
 
-data(:,1)=run(lag:length(run),index);
-data(:,2)=run(1:(length(run)-lag+1),index);
-
-ccolors = vector2colors(run(1:(length(run)-lag+1),2));
-scatter(data(:,1),data(:,2),5,ccolors,'filled'); 
-    
-
-hold on;
-
-for i=2:length(run)
-    
+for run=1:maxRuns
+    run
+    data = allRuns(find((allRuns(:,1)==run) & (allRuns(:,2)>2000)),6);
+    hursts(run) = estimate_hurst_exponent(data');
 end
 
-hold off;
+title('ACF for different values of D','FontSize',14);
+xlabel('Lag','FontSize',14);
+ylabel('Correlatoin','FontSize',14);
+legend('1','0.1','0.01','0.001','0.0001','0.00001');
+
+hold off
+
+ indicies=find(vec(:,7)==size);
+    fitx=(1+abs(vec(indicies,5)));
+    fity=(1+abs(vec(indicies,6)));
+
+    r=ksr(fitx,fity,1,100);
+
+    mins=10000000+zeros(length(r.x),1);
+    maxs=zeros(length(r.x),1);
+
+    for c=1:20
+        th=rand(length(fitx),1);
+        indicies=find(th(:)>0.95);
+        tfitx=fitx(indicies);
+        tfity=fity(indicies);
+        r=ksr(tfitx,tfity,1,100);
+        mins=min(mins,r.f');
+        maxs=max(maxs,r.f');
+    end
+
+    r=ksr(fitx,fity,1,100);
+
+    plot(r.x,r.f,'LineWidth',1,'Color',char(colors(size)));
+    ciplot(mins,maxs,r.x,char(colors(size)));
+
+    size
+
+filename = '..\\graphics\\comparisonOfACFs.pdf';
+orient tall;
+print('-dpdf', filename);

### 2008-05-08 documents/WCSS2008Paper/paper.tex
Adds explanation of Farmer deficiencies and graphs to paper.

@@
     \mbox{
-      \subfigure[Raw and absolute returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsPlot.pdf}}}
+      \subfigure[Raw and absolute log returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsPlot.pdf}}}
       \quad
-      \subfigure[Returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsHistogram.pdf}}}       
+      \subfigure[Distribution of Returns]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-returnsHistogram.pdf}}}       
       \quad
-      \subfigure[Log returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-logReturnsHistogram.pdf}}}       
+      \subfigure[Distribution of Returns (log scale)]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-s01-logReturnsHistogram.pdf}}}       
 }
@@
       \quad
-      \subfigure[Log returns histogram.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-logReturnsHistogram.pdf}}}       
+      \subfigure[Distribution of Returns (log scale)]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/Cont-logReturnsHistogram.pdf}}}       
 }
@@
 \begin{itemize}
-\item Patient agents place \texttt{LimitOrders} of constant size at a Poisson rate per unit time, and with a randomly generated price.
+\item Patient agents place \texttt{LimitOrders} of constant size $\sigma$ at a Poisson rate of $\alpha$ per unit time, and with a randomly generated price.
 Orders are equally likely to be either buy or sell orders.
-Buy limit orders are placed uniformly anywhere in the semiinfinite interval below the current ask price ($-\infty < p < a(t)$), where $p$ is the \emph{logarithm} of the price; similarly for sell orders.
+Buy limit orders are placed uniformly anywhere in the semi-infinite interval below the current ask price ($-\infty < p < a(t)$), where $p$ is the \emph{logarithm} of the price; similarly for sell orders.
 A uniform distribution in log prices is equivalent to an exponential distribution in prices.
-In addition, outstanding limit orders are canceled at a Poisson rate per unit time.
+In addition, outstanding limit orders are cancelled at a Poisson rate of $\delta$ per unit time.
 All of these processes are independent.
-\item Impatient agents place \texttt{MarketOrders} of constant size at a certain Poisson rate per unit time.
+\item Impatient agents place \texttt{MarketOrders} of constant size $\sigma$ at a Poisson rate of $\mu$ per unit time.
 \item A double auction order book.
@@
 \end{itemize}
-This model explains a large portion of the spread and price diffusion of actual stocks given an order flow rate, which essentially means that the double auction structure has a large impact on the nature of market movements.
+This model explains a large portion of the spread and price diffusion of actual stocks given an order flow rate, which implies that the double auction structure has a significant impact on the nature of market movements.
 The model also yields market behavior qualitatively close to actual markets in many respects, see Figure \ref{fig:sampleDynamicsFarmer}.
+However, as a model of market dynamics, it fails on some fronts.
+Notably, we do not see realistic volatility clustering.
+Volumes are fairly consistent, as are average returns.
+There is no lingering positive autocorrelation of absolute returns.
+Also notice the amplitude of single-period returns in Figure \ref{fig:sampleDynamicsFarmer}(g).
+The average per-step absolute returns are around $20\%$, much higher than empirically plausible, at least if a step is meant to signify short period, such as a day or less.
+This is clearly seen when we use actual prices in the order book.
+The same prices which look somewhat reasonable when viewed on a log scale appear quite volatile prices when viewed on a normal scale as in Figure \ref{fig:dynamicsFarmerNonLog}.
 
@@
    \mbox{
-      \subfigure[Ask and bid prices.]{\scalebox{0.5}{\quad \quad \includegraphics[width=1.0\textwidth]{../graphics/Farmer-pricesPlot.pdf}}} \quad
-      \subfigure[Raw and absolute returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnPlot.pdf}}}
+      \subfigure[Snapshot of Orderbook.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-orderbooks.pdf}}} 
+      \quad
+      \subfigure[Bid and ask log prices.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-pricesPlot.pdf}}}
 }
    \mbox{
-      \subfigure[Orderbook shape.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-orderbooks.pdf}}} \quad
+      \subfigure[Log returns histogram.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnsHistogram.pdf}}}
+      \quad
       \subfigure[Trade volumes.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-volumesPlot.pdf}}}
@@
 \mbox{
-      \subfigure[Returns histogram.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnsHistogram.pdf}}} \quad
+      \subfigure[Rolling Average Volatility.]{\scalebox{0.5}{\includegraphics[width=1.0\textwidth]{../graphics/Farmer-avgVolatility.pdf}}}
+      \quad
       \subfigure[Autocorrelation of returns.]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-acfReturns.pdf}}}
       }
-    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation for default Farmer's parametrization.}
+\mbox{
+      \subfigure[Raw and absolute log returns.]{\scalebox{0.6}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-returnPlot.pdf}}}
+     }
+    \caption{Examples of outputs and statistics from a single run of the Farmer FinancialModel simulation with $\alpha=.585$, $\delta=.39$, $\mu=.325$, $\sigma=1$.
+These parameters are derived from empirical data.}
     \label{fig:sampleDynamicsFarmer}
+  \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Snapshot of Orderbook with actual prices (log x scale).]{\scalebox{0.6}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-linear-orderbooks.pdf}}} 
+}
+   \mbox{
+\subfigure[Prices (log scale).]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-linear-pricesPlot-logscale.pdf}}}
+      \quad
+\subfigure[Same prices (natural scale).]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/Farmer-linear-pricesPlot.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation with $\alpha=.585$, $\delta=.39$, $\mu=.325$, $\sigma=1$.
+These parameters are derived from empirical data.}
+    \label{fig:dynamicsFarmerNonLog}
   \end{center}
@@
    \mbox{
-      \subfigure[Autocorrelation functions]{\scalebox{0.5}{\quad \quad \includegraphics[width=1.0\textwidth]{../graphics/comparisonOfACFs.pdf}}} \quad
-      \subfigure[Return distributions]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth]{../graphics/comparisonOfHistograms.pdf}}}
+      \subfigure[Autocorrelation functions]{\scalebox{0.45}{\quad \quad \includegraphics[width=1.0\textwidth]{../graphics/comparisonOfACFs.pdf}}} \quad
+      \subfigure[Return distributions]{\scalebox{0.45}{\includegraphics[width=1.00\textwidth]{../graphics/comparisonOfHistograms.pdf}}}
       }

### 2008-05-08 src/gui/FinancialModelWithUI.java
Minor improvement and added comments

@@ -148,22 +148,18 @@ public class FinancialModelWithUI extends GUIState {
 						myReporter.updateTrailing();
 					}
 
-					// TODO: why set series three times??
-					if (priceFrame.isVisible()) {
+					if (priceFrame.isVisible() || 
+						returnFrame.isVisible() ||
+						volumeFrame.isVisible()) 
+					{
 						priceChart.disable();
-						myReporter.setSeries();
-						priceChart.enable();
-					}
-
-					if (returnFrame.isVisible()) {
 						returnChart.disable();
-						myReporter.setSeries();
-						returnChart.enable();
-					}
-
-					if (volumeFrame.isVisible()) {
 						volumeChart.disable();
+
 						myReporter.setSeries();
+					
+						priceChart.enable();
+						returnChart.enable();
 						volumeChart.enable();
 					}
 
@@ -280,8 +276,8 @@ public class FinancialModelWithUI extends GUIState {
 		orderHist.setRangeAxisLabel("Number of observations");
 		orderHist.getChartPanel().getChart().setBackgroundPaint(Color.white);
 		for (int a = 0; a < myModel.parameterMap.get("numAssets"); a++) {
-			orderHist.addSeries(fakeArray, 150, "sell orders for asset " + a, null);
 			orderHist.addSeries(fakeArray, 150, "buy orders for asset " + a, null);
+			orderHist.addSeries(fakeArray, 150, "sell orders for asset " + a, null);
 		}
 		orderHist.update();
 		orderHistFrame = orderHist.createFrame(this);

### 2008-05-08 src/gui/GUIReporter.java

@@ -39,12 +39,14 @@ public class GUIReporter implements Steppable {
 	
 	public ArrayList<Vector<Double>> volumeMemory = new ArrayList<Vector<Double>>();
 	
-	public int NumViewable = 10000; // also, the #/data points used in
+	// The number of data points saved and shown for the price, return, and volume series'
+	// Also, the #/data points used in calculating autocorrelation
+	public int NumViewable = 10000; 
 
-	// calculating acf
-
+	// number of steps "lag" to compute the autocorrelation for
 	public int ACFViewable = 200;
 
+	// number of steps to average over when doing trailing averages
 	public int trailingSampleSize = 365;
 	
 	FinancialModel myModel;

### 2008-05-08 documents/WCSS2008Paper/paper.tex
Added Hurst exponent exploration for Cont to the paper.

@@
       }
-    \caption{Autocorrelation and distributions in Cont simulation with $s=0.1$ and varying $D$, averaged over many runs.}
+
+      \subfigure[Hurst exponents for different $s$]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth, angle=270]{../graphics/comparisonOfHurstsS}}} 
+      \subfigure[Hurst exponents for different $D$]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth, angle=270]{../graphics/comparisonOfHurstsD}}}
+
+    \caption{Autocorrelation and distributions in Cont simulation with $s=0.1$ and varying $D$, averaged over many runs.
+In bottom row, kernel regresion on Hurst exponents of time series generated from Con't model for varying levels of $s$ and $D$ is presented.}
     \label{fig:ContMultiRun}

### 2008-05-08 scripts/plotComplex.m

@@ -1,51 +1,55 @@
 filename='..\\timeSeries.txt';
 [h, allRuns]=hdrload(filename);
 
-maxRuns=100;
+maxRuns=190;
 
 hursts=zeros(maxRuns,1);
-
-hold all;
+inputs=zeros(maxRuns,1);
 
 for run=1:maxRuns
     run
-    data = allRuns(find((allRuns(:,1)==run) & (allRuns(:,2)>2000)),6);
-    hursts(run) = estimate_hurst_exponent(data');
+    data = allRuns(find((allRuns(:,1)==run) & (allRuns(:,2)>2000)),:);
+    hursts(run) = estimate_hurst_exponent(data(:,6)');
+    inputs(run)=mean(data(:,3));
 end
 
-title('ACF for different values of D','FontSize',14);
-xlabel('Lag','FontSize',14);
-ylabel('Correlatoin','FontSize',14);
-legend('1','0.1','0.01','0.001','0.0001','0.00001');
 
-hold off
 
- indicies=find(vec(:,7)==size);
-    fitx=(1+abs(vec(indicies,5)));
-    fity=(1+abs(vec(indicies,6)));
 
-    r=ksr(fitx,fity,1,100);
+points=100;
+spacing=0.01;
 
-    mins=10000000+zeros(length(r.x),1);
-    maxs=zeros(length(r.x),1);
+fitx=inputs;
+fity=hursts;
 
-    for c=1:20
-        th=rand(length(fitx),1);
-        indicies=find(th(:)>0.95);
-        tfitx=fitx(indicies);
-        tfity=fity(indicies);
-        r=ksr(tfitx,tfity,1,100);
-        mins=min(mins,r.f');
-        maxs=max(maxs,r.f');
-    end
+r=ksr(fitx,fity,spacing,points);
 
-    r=ksr(fitx,fity,1,100);
+mins=10000000+zeros(length(r.x),1);
+maxs=zeros(length(r.x),1);
 
-    plot(r.x,r.f,'LineWidth',1,'Color',char(colors(size)));
-    ciplot(mins,maxs,r.x,char(colors(size)));
+for c=1:20
+    th=rand(length(fitx),1);
+    indicies=find(th(:)>0.2);
+    tfitx=fitx(indicies);
+    tfity=fity(indicies);
+    r=ksr(tfitx,tfity,spacing,points);
+    mins=min(mins,r.f');
+    maxs=max(maxs,r.f');
+end
 
-    size
[…]

### 2008-05-08 documents/WCSS2008Paper/paper.tex

@@
 
-TODO Note: liquidity issues
-
 \section{Verification of Correctness}
 
-TODO: show results side by side with original papers' figures
-
+The supreme objective was to recreate results reported by others rather than to create a standalone model of the market.
+Therefore, the validation focused on comparing outputs from our simulations to graphs and statistics from other papers.
 These compare well with data shown in the original paper (shown in Figures \ref{fig:ContSmallSPap} and \ref{fig:ContLargeSPap}).
@@
     \caption{Autocorrelation and distributions in Cont simulation with $s=0.1$ and varying $D$, averaged over many runs.
-In bottom row, kernel regresion on Hurst exponents of time series generated from Con't model for varying levels of $s$ and $D$ is presented.}
+In bottom row, kernel regresion with 95\% resampled confidence intervals on Hurst exponents of time series generated from Con't model for varying levels of $s$ and $D$ is presented.}
     \label{fig:ContMultiRun}

### 2008-05-08 scripts/plotComplex.m

@@ -43,13 +43,13 @@ ciplot(mins,maxs,r.x,'red');
 plot(r.x,r.f,'LineWidth',2,'Color','red');
 hold off;
 
-title('Hurst exponent for different values of s','FontSize',14);
-xlabel('Value of s','FontSize',14);
+title('Hurst exponent for different values of D','FontSize',14);
+xlabel('Value of D','FontSize',14);
 ylabel('Hurst exponent','FontSize',14);
 
 alpha(0.5)
     
-filename = 'comparisonOfHurstsS.bmp';
+filename = 'comparisonOfHurstsD.bmp';
 orient landscape;
 print('-dbmp', filename);
 

### 2008-05-08 documents/WCSS2008Paper/paper.tex

@@
 
+\subsection{Westerhoff Model}
+
+The Westerhoff model (\cite{westerhoff2004}) includes a limited number of markets in which two types of agents operate.
+Fundamentalists specialize in a certain market while chartists may switch between markets.
+More specifically, chartists use extrapolative methods to select markets that shows price trends but that are not too misaligned.
+
+The model considers $k = 1,2\dots K$ asset markets of equal size: the log fundamental value of asset $k$ in period $t + 1$ is regulated by a news arrival process:
+\begin{equation*}
+F_{t+1}^k = F_t^k + N
+\end{equation*}
+
+Market makers are present in all markets and stand ready to absorb imbalances between buyers and sellers: they adjust prices $S^k_t$ according to a loglinear price impact function.
+Traders submit orders according to their expectations of the price movements. Fundamentalist expect the prices of the assets to return to their fundamental values and their demand is expressed as:
+\begin{align*}
+D_t^{F,k} &= a^F[E_t^F(S^k_{t+1}) - S^k_{t}],\\
+E_t^F &= S^k_{t} + b^F(F_t^k - S^k_t)
+\end{align*}
+On the other hand, chartists follow simple technical analysis rules:
+\begin{align*}
+D_t^{C,k} &= a^C[E_t^C(S^k_{t+1}) - S^k_{t}],\\
+E_t^C &= S^k_{t} + b^C(S_t^k - S^k_{t-1})
+\end{align*}
+
+Chartists try to identify the attractiveness of markets with a fitness measure that entails the risk of being captured in a bursting bubble. The attractiveness of market $k$ is defined as:
+\begin{equation*}
+A^k_t = \log \frac{1}{1+f(F_t^k - S^k_t)^2}
+\end{equation*}
+The relative percentage of chartists choosing market $k$ depends on the attractiveness of that market and is defined as:
+\begin{equation*}
+W^k_t = \frac{\exp(gA^k_t)}{\sum_{k=1}^K \exp(gA^k_t)}
+\end{equation*}
+As shown in Figure \ref{fig:Westerhoffsim}, this model is capable of  generating complex price dynamics, volatility clustering, fat tails for the distribution of the returns and positive autocorrelation for absolute returns.
+\begin{figure}[htbp]
+  \begin{center}
+   \mbox{
+      \subfigure[Log price $S^1$]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/s1dyn.pdf}}}
+      \quad
+      \subfigure[Log price index]{\scalebox{0.33}{ \includegraphics[width=1.0\textwidth]{../graphics/Idyn.pdf}}} \quad
+      \subfigure[Fraction of chartist in market 1]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/w1dyn.pdf}}}
+      }
+    \mbox{
+      \subfigure[Raw returns of asset 1]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/rets1.pdf}}}
+      \quad
+      \subfigure[Raw returns of price index ]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/retI.pdf}}}
+      \quad
+       \subfigure[Attractor for log price changes of asset 1]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/phasespaceS1.pdf}}}   
+   }
+    \mbox{
+      \subfigure[Attractor for price index changes]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/phasespaceI.pdf}}}
+      \quad
+      \subfigure[Autocorrelation for raw returns of asset 1]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/acfS1.pdf}}}
+      \quad
+      \subfigure[Autocorrelation of absolute returns of asset 1]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/acfabsS1.pdf}}}
+      }
+    \caption{Examples of outputs and statistics from a single run of the Westerhoff simulation with $K = 5$, $N=0.0002$, $a^M=0.1$, $a^c, b^c = 5$, $a^F, b^F = 0.2$, $f = 1,000,000$ and $g = 1.2$.}
+    \label{fig:Westerhoffsim}
+  \end{center}
+\end{figure}
+
 \section{Verification of Correctness}
@@
 
-      \subfigure[Hurst exponents for different $s$]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth, angle=270]{../graphics/comparisonOfHurstsS}}} 
-      \subfigure[Hurst exponents for different $D$]{\scalebox{0.5}{\includegraphics[width=1.00\textwidth, angle=270]{../graphics/comparisonOfHurstsD}}}
+      \subfigure[Hurst exponents for different $s$]{\scalebox{0.5}{\includegraphics[width=0.9\textwidth, angle=270]{../graphics/comparisonOfHurstsS}}} 
+      \subfigure[Hurst exponents for different $D$]{\scalebox{0.5}{\includegraphics[width=0.9\textwidth, angle=270]{../graphics/comparisonOfHurstsD}}}
 

### 2008-05-08 documents/WCSS2008Paper/paper.tex
Adds more description for Farmer-Cont and graphs with a more flattering parameterization

@@
 
+\paragraph*{}
 One down side of the Cont model that is not noted in the papers is that there appears to be a great degree of periodic cyclicality in the volatility of the markets.
@@
 \end{itemize}
+\paragraph*{}
 This model explains a large portion of the spread and price diffusion of actual stocks given an order flow rate, which implies that the double auction structure has a significant impact on the nature of market movements.
 The model also yields market behavior qualitatively close to actual markets in many respects, see Figure \ref{fig:sampleDynamicsFarmer}.
+\paragraph*{}
 However, as a model of market dynamics, it fails on some fronts.
@@
 Impatient traders are modeled after Cont.
-They have heterogeneous volatility thresholds, and place market orders depending upon how public information in the form of a normally distributed random variable compares with these thresholds.
+They have heterogeneous volatility thresholds, and place market orders depending upon how public information in the form of a normally distributed random variable $\epsilon_t ~ N(0,D^2)$ compares with these thresholds.
 Some fraction $s$ of agents then adjust their volatility thresholds in light of current volatility by setting it to the trailing period's return rate.
+
+\paragraph*{}
+The result is shown in Figure \ref{fig:FarmerContDynamics}.
+When there are sufficient limit orders being placed by Farmer's patient agents to ensure adequate liquidity for Market orders, we get dynamics that have many positive aspects compared with reality --- approximately power law return distribution, volatility clustering, realistic variability in linear prices, volumes, and returns, and essentially no predictability of future returns based upon past returns.
 
@@
       \quad
-      \subfigure[Autocorrelation of absolute returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-acfReturns.pdf}}}
+      \subfigure[Autocorrelation of returns.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-acfReturns.pdf}}}
       }
-    \caption{Examples of outputs and statistics from a single run of the Farmer-Cont hybrid.}
-    \label{fig:ContLargeSSim}
+    \caption{Examples of outputs and statistics from a single run of the Farmer-Cont hybrid FinancialModel simulation with actual prices, $D=0.001$, $s=0.01$, $\alpha=0.9$, $\delta=0.2$, and $\mu=0.4$.}
+    \label{fig:FarmerContDynamics}
   \end{center}

### 2008-05-08 documents/WCSS2008Paper/paper.tex
Recompiles paper with latest graphs.. also minor fixes to latex code.

@@
 They are heterogeneous in the sense that they are given independently assigned (subjective) volatility thresholds $\theta_i(t)$.
-Each period, all agents receive a common signal, which can be interpreted as public information or "news", in the form of IID Gaussian random variables $\epsilon_t ~ N(0,D^2)$.
+Each period, all agents receive a common signal, which can be interpreted as public information or "news", in the form of IID Gaussian random variables $\epsilon_t \sim N(0,D^2)$.
 Each agent $i$ responds to this signal by selling if $\epsilon_t < -\theta_i(t)$, buying if $\epsilon_t > \theta_i(t)$, and otherwise sitting out the period.
@@
       }
-    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation with $\alpha=.585$, $\delta=.39$, $\mu=.325$, $\sigma=1$.
-These parameters are derived from empirical data.}
+    \caption{Examples of outputs and statistics from a single run of the FinancialModel simulation with $\alpha=.585$, $\delta=.39$, $\mu=.325$, $\sigma=1$ shown in actual, rather than log prices}
     \label{fig:dynamicsFarmerNonLog}
@@
 Impatient traders are modeled after Cont.
-They have heterogeneous volatility thresholds, and place market orders depending upon how public information in the form of a normally distributed random variable $\epsilon_t ~ N(0,D^2)$ compares with these thresholds.
+They have heterogeneous volatility thresholds, and place market orders depending upon how public information in the form of a normally distributed random variable $\epsilon_t \sim N(0,D^2)$ compares with these thresholds.
 Some fraction $s$ of agents then adjust their volatility thresholds in light of current volatility by setting it to the trailing period's return rate.
@@
 The result is shown in Figure \ref{fig:FarmerContDynamics}.
-When there are sufficient limit orders being placed by Farmer's patient agents to ensure adequate liquidity for Market orders, we get dynamics that have many positive aspects compared with reality --- approximately power law return distribution, volatility clustering, realistic variability in linear prices, volumes, and returns, and essentially no predictability of future returns based upon past returns.
+When there are sufficient limit orders being placed by Farmer's patient agents to ensure adequate liquidity for Market orders, we get dynamics that have many positive aspects compared with reality --- approximately power law return distribution, volatility clustering, realistic variability in \emph{actual} prices, volumes, and returns, and essentially no predictability of future returns based upon past returns.
 
@@
       \quad
-      \subfigure[Snapshot of orderbook.]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-orderbooks.pdf}}}
+      \subfigure[Snapshot of orderbook (log x scale).]{\scalebox{0.33}{\includegraphics[width=1.00\textwidth]{../graphics/FC-orderbooks.pdf}}}
       \quad
